%!TEX root = ../thesis.tex
% the acknowledgments section

This thesis is the culmination of four years of undergraduate studies and several months of research. I would not have gotten to this point without the constant support of many individuals.

I would like to express my deep gratitude to Professor Moon Duchin, my thesis advisor, but also my mentor and teacher for the past two years. She inspired me, encouraged me, taught me, and believed in me. Without her this thesis would have been impossible.

I would like to thank my mom for her unwavering love and selflessness. She has given so much of herself to ensure that I was able to follow my dreams, and for that I am immensely thankful. Although I am far from her and home, her care propels me forward.

I would like to thank Dr. Thomas Weighill, whose patience, ingenuity, generosity, and brilliance contributed greatly to this research.

I would like to thank the researchers of the Metric Geometry and Gerrymandering Group, particularly Ruth Buck and Dr. Daryl DeFord. They were instrumental in my early understandings of computational research on gerrymandering, and gave me opportunities to grow.

I would like to thank Professor Latanya Sweeney and Jinyan Zang. They provided me with patient guidance, enthusiastic encouragement and useful feedback on this research.

Finally, I would like to thank my classmates and friends. They have toiled beside me at all hours, listened to me, helped me and held me throughout this experience. Thank you.
